Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-05-17 ms.tex
Adding three panel filler plot until replaced
Adding color coding for sections that are under construction

@@

Nice summary paragraph from recent Bourier  papers.

@@

\citet{1973MNRAS.162...43H, 1974MNRAS.166..373H} started work on analytic solutions as you get a Laplace equation. Doesn't get boundary condition right.
\citet{1990ApJ...350..216N} has analytic solutions following Harrington.
\citet{1990ApJ...350..216N} improved on Harrington. Does have divergent inhomogeneous solution in slab geometry! Also has discussion of absorption.
\citet{2006ApJ...649...14D} generalized Harrington to spherical geometry
\citet{1976ApJ...208..286W} tries to compute the radiation force
\citet{1994ApJ...427..603R} discuss an ansatz for the fokker-planck approximation where the full voigt profile is used.
\citet{2020arXiv200509692L} discusses solutions for the sphere. Follows \citet{2006ApJ...649...14D}. They cite \citet{2015MNRAS.449.4336S} on core skipping.
\citet{2002ApJ...567..922A,2015MNRAS.449.4336S} discusses acceleration.
\citet{2015MNRAS.449.4336S}: amounts to drawing the atom velocity from a truncated Gaussian distribution to force it back into the wing.  The logic seems to be that almost all of the core scattering results in zero spatial diffusion until you get far enough from the line wing.  It seems like this is just calibrated by modifying the truncation until the spectrum doesn’t seem to depend sensitively on the choice of truncation.

\section{ no photon destruction }
\label{sec:no_destruction}

@@

TODO: Add Monte Carlo description

\subsection{Results}
\label{subsec:no_destruction:results}
\begin{figure}
    \centering
    \includegraphics{pdf_xinit0.0.pdf}
    \caption{Analytic solution for the outgoing spectrum compared with Monte Carlo at a line center optical depth of $\tau_0 = 1 \times 10^7$. Photons are initialized with $\rm x = 0.0$. In the log-scale plot in the second panel, $|\rm H_{bc}|$ is shown instead of $\rm H_{bc}$. Residuals of each solution to the Monte Carlo data points are shown in the third panel.} 
    \label{fig:sol_mc_residual_0}
\end{figure}

@@

 TODO: Add optical depth comparison figure, $x_s$ comparison figure

\section{Time-Dependent Diffusion}
\label{sec:time_dependent}

@@

\subsection{Numerical Methods}
\label{subsec:time_dependent:numerical_methods}

@@

(TODO: Rewrite) Here we have closed the integral along the $\gamma$ axis by a semi-circle at infinity, using the residue theorem to obtain the answer. The contribution from the semi-circle at infinity is zero. Summing over all spatial eigenmodes $n$ and over all temporal eigenmodes $m$ for a given $n$, we get the final form of the result
\be
J(r,\sigma,t) & = & \sum_{n=1}^\infty j_0(\kappa_n r)  \sum_m J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Using the derivative
\be
\frac{dj_0(\kappa_n r)}{dr} \rfloor_R & =& \frac{d}{dr}  \rfloor_R \left( \frac{\sin(\kappa_n r)}{\kappa_n r} \right)
=  \left( \frac{\cos(\kappa_n R)}{R} - \frac{\sin(\kappa_n R)}{\kappa_n R^2} \right)\rfloor_R
\nonumber \\ & = & \frac{(-1)^n}{R},
\ee
the flux at the surface is
\be
F(R,\sigma,t) & =& - \frac{4\pi}{3k\phi} \frac{dJ(R,\sigma,t)}{dr} 
= - \frac{4\pi}{3k\phi R}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Multiplying by $4\pi R^2$ gives the energy per time per frequency
\be
\frac{dE}{dtd\nu} & = & - \frac{16\pi^2 R}{3k\phi}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\label{eq:dEdtdnu}
\ee
Integrating this over time yields a factor $-1/\gamma_{nm}$, and over $d\nu = \sqrt{3/2} \Delta^2 \phi d\sigma$, a ``sum rule" results
\be
1 & = &  \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE} \sum_{nm} (-1)^n \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
This non-trivial expression provides a check on the values of $\gamma_{nm}$ and the functions $J_{nm}(\sigma)$. This expression can also be written as
\be
1 & =& \sum_{nm} P_{nm},
\label{eq:sumrule}
\ee
where the contribution of each eigenfunction is
\be
P_{nm} & \equiv & \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE}  (-1)^n \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
These coefficients can have either sign depending on the value of $n$.

>>>>

%%% FOR DRAFTING: REMOVE AFTER PAPER IS COMPLETE!
\usepackage{xcolor}

@@

\textcolor{red}{Nice summary paragraph from recent Bourier  papers.}

@@

\citet{1973MNRAS.162...43H, 1974MNRAS.166..373H} demonstrated that resonance-line radiation transfer in low-density, optically-thick media can be described by the Poisson equation. However, the accuracy of their solution is limited by the treatment of the boundary conditions. In their eigenfunction expansion, the spatial and frequency components of the differential equation are separated, yet the spatial eigenvalues are defined as being explicitly frequency-dependent. The resulting solution for large optical depths would suggest the mean intensity at the boundary of the slab is zero, which is unphysical. A more detailed discussion of the boundary condition is included in Section \ref{sec:no_destruction}. \citet{1990ApJ...350..216N} follows this approach in an attempt to extend the result to media of intermediate optical depth, including the effects of scattering in the Doppler core of the line, but again does not use the correct boundary condition. \citet{1990ApJ...350..216N} improved on this approach, producing a divergent, inhomogeneous solution to the problem in 1D slab geometry that includes the effect of photon absorption within the medium. \citet{2006ApJ...649...14D} generalize the problem to spherical geometry, which is the same physical set up used here. Following this, \citet{2020arXiv200509692L} discusses solutions for the sphere, utilizing the ``core skipping'' method discussed in \citet{2015MNRAS.449.4336S} to handle photon frequency redistribution within the Doppler core.

%%%%%%%%% TEMPORARY %%%%%%%%%%%%%%%
\textcolor{red}{
\citet{1973MNRAS.162...43H, 1974MNRAS.166..373H} started work on analytic solutions as you get a Laplace equation. Doesn't get boundary condition right.
\citet{1990ApJ...350..216N} has analytic solutions following Harrington.
\citet{1990ApJ...350..216N} improved on Harrington. Does have divergent inhomogeneous solution in slab geometry! Also has discussion of absorption.
\citet{2006ApJ...649...14D} generalized Harrington to spherical geometry
\citet{1976ApJ...208..286W} tries to compute the radiation force
\citet{1994ApJ...427..603R} discuss an ansatz for the fokker-planck approximation where the full voigt profile is used.
\citet{2020arXiv200509692L} discusses solutions for the sphere. Follows \citet{2006ApJ...649...14D}. They cite \citet{2015MNRAS.449.4336S} on core skipping.
\citet{2002ApJ...567..922A,2015MNRAS.449.4336S} discusses acceleration.
\citet{2015MNRAS.449.4336S}: amounts to drawing the atom velocity from a truncated Gaussian distribution to force it back into the wing.  The logic seems to be that almost all of the core scattering results in zero spatial diffusion until you get far enough from the line wing.  It seems like this is just calibrated by modifying the truncation until the spectrum doesn’t seem to depend sensitively on the choice of truncation.  
}
\section{ no photon destruction }
\label{sec:no_destruction}

@@

\textcolor{red}{
TODO: Add Monte Carlo description
}
\subsection{Results}
\label{subsec:no_destruction:results}
\begin{figure}
    \centering
    \includegraphics{pdf_xinit0.0.pdf}
    \caption{Analytic solution for the outgoing spectrum compared with Monte Carlo at a line center optical depth of $\tau_0 = 1 \times 10^7$. Photons are initialized with $\rm x = 0.0$. In the log-scale plot in the second panel, $|\rm H_{bc}|$ is shown instead of $\rm H_{bc}$. Residuals of each solution to the Monte Carlo data points are shown in the third panel.} 
    \label{fig:sol_mc_residual_0}
\end{figure}

@@

 \begin{figure}
    \centering
    \includegraphics{threepanel_test.png}
    \caption{Analytic solution for the outgoing spectrum compared with Monte Carlo at a line center optical depth of $\tau_0 = 1 \times 10^7$. Photons are initialized with $\rm x = 0.0, 6.0, 12.0$ to examine the solutions' accuracy with a changing source frequency. TODO: Remove blue line. We really don't care about the divergent solution anymore.} 
    \label{fig:sol_mc_xinit}
\end{figure}

 \textcolor{red}{
 TODO: Add optical depth comparison figure, $x_s$ comparison figure
}
\section{Time-Dependent Diffusion}
\label{sec:time_dependent}

@@

%%% TODO: NOTES
% Can't solve this exactly, need a numerical scheme to solve. Now going to turn to solving this equation (without subsec). We start with a treatment of the boundary conditions. Continuity in J at the source, the discontinuity needs to be (Eq. 24)

% Solving the sigma dependence
% Solving the time dependence
% Solving the frequency dependence

% Get different pieces of the solution and build it up.

% Make sure that all variables that are introduced in the text are defined beforehand. Especially those that are defined only in the appendix.

@@

\textcolor{red}{(TODO: Rewrite) Here we have closed the integral along the $\gamma$ axis by a semi-circle at infinity, using the residue theorem to obtain the answer. The contribution from the semi-circle at infinity is zero. Summing over all spatial eigenmodes $n$ and over all temporal eigenmodes $m$ for a given $n$, we get the final form of the result}
\be
J(r,\sigma,t) & = & \sum_{n=1}^\infty j_0(\kappa_n r)  \sum_m J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Using the derivative
\be
\frac{dj_0(\kappa_n r)}{dr} \rfloor_R & =& \frac{d}{dr}  \rfloor_R \left( \frac{\sin(\kappa_n r)}{\kappa_n r} \right)
=  \left( \frac{\cos(\kappa_n R)}{R} - \frac{\sin(\kappa_n R)}{\kappa_n R^2} \right)\rfloor_R
\nonumber \\ & = & \frac{(-1)^n}{R},
\ee
the flux at the surface is
\be
F(R,\sigma,t) & =& - \frac{4\pi}{3k\phi} \frac{dJ(R,\sigma,t)}{dr} 
= - \frac{4\pi}{3k\phi R}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Multiplying by $4\pi R^2$ gives the energy per time per frequency
\be
\frac{dE}{dtd\nu} & = & - \frac{16\pi^2 R}{3k\phi}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\label{eq:dEdtdnu}
\ee
Integrating this over time yields a factor $-1/\gamma_{nm}$, and over $d\nu = \sqrt{3/2} \Delta^2 \phi d\sigma$, a ``sum rule" results
\be
1 & = &  \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE} \sum_{nm} (-1)^n \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
This non-trivial expression provides a check on the values of $\gamma_{nm}$ and the functions $J_{nm}(\sigma)$. This expression can also be written as
\be
1 & =& \sum_{nm} P_{nm},
\label{eq:sumrule}
\ee
where the contribution of each eigenfunction is
\be
P_{nm} & \equiv & \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE}  (-1)^n \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
These coefficients can have either sign depending on the value of $n$.
